### 2011-10-19 doc/manuscript.tex
I am editting the Results section that had been modified by Ilan.

\section*{A list our major and minor findings or observations}
\begin{enumerate}
\item We showed that recombining gene transfer from SPY to SDE was strong
enough.
\item A few evidences for the net directionality from SPY to SDE:
time-constrained gene transfer between SPY to SDE also showed the directionality
from SPY to SDE.
\item We showed that virulence genes were associated with horizontal gene
transfer. 
\item We found that recombining gene tranfer processes are involved in
membrane-bound protein encoding genes and aminoacyl tRNA synthetases.
\item We found that horizontal gene transfer processes are involved in
transposases.
\item We provided evidence that recombination process affected more sites along
the core genome of SPY and SDE than mutation process.
\item We found a recombination hotspot at 200 kb from the origin of replication
along the genome of SDE1.
\item We showed that species-tree-topology-changing recombinant edges were more
reliably recovered than species-tree-topology-not-changing edges.
\item We showed that recombination intensity might be useful in finding genes
under high recombining gene transfer.
\item We found that SPY and SDE are more closely related than SDD and SDE in
about 10\% of the core genome.
\item Recombination intensity and number of recombining gene transfer events
were positively correlated.
\item We found reduction of numbers of genes after divergence of SDE and SPY.
\item We confirmed the species tree that relates the five genomes.
\item We listed several virulence genes associated with recombining and
horizontal gene transfer.
\end{enumerate}

\section*{Suggestions about the manuscript}
\begin{enumerate}
\item Ilan thought that a species tree must be shown at the outset of the
section \textbf{Results}.
\item Ilan suggested a different order of four parts in \textbf{Results}:
preliminary, funtional category and virulence gene analyses, gene gain and loss,
and directionality.
\end{enumerate}

@@

Therefore, the research topic on gene exchange between SPY and SDE has been
performed in various ways. Sachse et al.\ \citet{Sachse2002} showed gene transfer from SDE to
SPY for the downstream of the genomic region of streptococcal pyogenic
exotoxins. Davies et al.\ \citet{Davies2007} investigated phage 3396 from SDE as a vehicle of
gene transfer from SDE to SPY. They also \citet{Davies2009} demonstrated that an
integrative conjugative element of SDE allowed gene transfer from SDE to other
streptococci. Ahmad et al.\ \citet{Ahmad2009} did not conclusively deduce the net
directionality of housekeeping gene transfer between SPY and SDE.  Notably,
Kalia et al.\ \citet{Kalia2001} attempted to show that the net direction of gene transfer from
SPY to SDE was more dominant than the other direction.  Because these researches
on the controversial issue, net directionality of gene transfer between SPY and
SDE, have been limited to relatively small regions in the genomes primarily on
representative loci from the Multi Locus Sequence Typing (MLST) database, it
remains to be elusive to determine the genome-wide degrees to which genomic
segments flowed between the two species, and net directions of thereof.

As sequencing costs plummet \citep{Mardis2011}, bacterial genome sequencing
produces individual genomes of high quality even within a bacterial species
\citep{Tettelin2009a}.  Studies on horizontal gene transfer in prokaryotes have
used genes in very large sets of species \citep{Koonin2001} from a great deal of
genomes available.  Although this type of study gave us insights into deep
evolutionary histories of prokaryotes, with multiple individual genomes per
species we now can go beyond such rough broad study on gene transfer towards
fine-scaled history of gene transfer between very closely related prokaryotes.
For instance, Liu et al.\ \citet{Liu2009} studied gene transfer events in
\textit{Lactobacillus bulgaricus} and \textit{Streptococcus thermophilus} using
the compositional approach of Karlin \citet{Karlin2001}. Luo et al.\ \citet{Luo2011} investigated
gene transfer events in \textit{Eschericha Coli} using the embedded quartet
decomposition due to Zhaxybayeva et al.\ \citet{Zhaxybayeva2006}. Hamady et al.\ \citet{Hamady2006} combined
phylogenetic and compositional approaches to detecting horizontal gene transfer.
Caro-Quintero et al.\ \citet{Caro-Quintero2011} studied a number of \textit{Shewanella baltica}
genomes to identify the genetic elements that enabled the species to adapt to
redox gradients.  Although several complete genomes from SPY have been
available, Suzuki et al.\ \citet{Suzuki2011} just recently added one more complete genome of
SDE to the set of a previously unique SDE genome \citep{Shimomura2011}. These
multiple SDE genomes prompted us to study genome-wide gene flows between SPY and
SDE that afflicted human beings around the globe.

@@

Recently, a method of inferring
homologous recombination with whole bacterial genomes was developed
\citep{Didelot2010}. The model of homologous recombination allowed them to study
flows of genetic segments between bacteria individuals of \textit{Bacillus
cereus} species in the core genomes.   
% Descriptions of ClonalOrigin model.
Other evolutionary processes acting on
bacterial genomes include genome rearragement, translocation, inversion,
duplication, gain, loss, etc.  These processes did not readily lend themselves
to statistically more sound model-based approaches such as maximum likelihood.
Instead, a parsimony-based approach to inferring gene duplication, loss, and
transfer events was developed by Doyon et al.\ \citet{Doyon2011}.  
% Descriptions of Mowgli analsysis.
We noted that the first
approach could be used to infer transfer of genomic segments due to homologous
recombination, and the second could be useful in studying horizontal gene
transfer.
% Homologous recombining gene transfer could be studied more readily using the
% core genomes of individual bacteria while horizontal gene transfer might be more
% easily detected using the pan genomes.

In short, the complete genomes of the two species of SPY and SDE presented us a
unique opportunity to study gene transfer over relatively recent time scales.
We took advantage of the model-based approach for inferring homologous
recombination, and the parsimony-based one for detecting gene duplication, loss,
and horizontal transfer. Virulence genes were found to be more associated
non-recombining horizontal gene transfer than non-virulence genes.  We compared
both of the results in recombining gene transfer events to show that the
parsimony-based method inferred recombining gene transfer events in biological
functional categories similar to those inferred with the model-based approach.
This was the first attempt as far as we know of systematically analyzing SPY and
SDE in genome-wide scale to detect gene transfer events not only in the core
genome but also in the pan genome, which made the analysis complete in terms of
the amount of used genome data. Our analysis of the core genomes also showed for the
first time that the net directionality of recombining gene transfer from SPY to
SDE was stronger at the genome scale.

@@

\subsection*{Distinguishing two gene transfer processes shows different functional
categories associations}

@@

\subsection*{Recombining gene transfer from SPY to SDE at genome-wide scale}

We show a net directionality of recombining gene transfers from SPY to SDE.

Describe our arguments of the net directionality. 
Number of recombinant edges.
Confidence about the numbers of recombinant edges.
Not all of genes experience recombining gene transfers from SPY to SDE.
Gene transfer counts from the parsimony-based approach; the power of the
parsimony-based approach may be limited.
Is this number of recombinant edges good enough for the argument?

Reflection of the previous analyses, 
heat maps of real data analysis, 
and supporting simulations.

\subsection{Bayesian analysis of recombining gene transfer}

{\bf IG: REMOVED A LOT OF DETAIL (MAINLY IN FIRST PART). PLEASE CHECK.}

We followed the pipeline and methods suggested in \cite{Didelot2010} to pursue a model based study of recombining gene transfer. A preliminary step in this analysis involved obtaining an estimate of the species tree with branch lengths (via ClonalFrame), and estimates of the three population parameters: mutation rate, recombination rate, and average recombinant tract length (via a first phase of ClonalOrigin). Due to divergence of the SEE sample from the other five, we could not obtain suficiently long reliable multiple sequence alginments with it, and had to exclude it from the bayesian analysis. The spcies tree inferred by ClonalFrame over the five species has the expected topology (Fig. {\bf [ADD FIGURE 1]} (see Figure S\ref{fig:tree5} for branch length estimates). ClonalFrame also produces estimates for the relative strength of recombination compared
to mutations ($r/m$; Table S\ref{tab:clonalframe}).  The estimate of $r/m=4.66$ provides an initial indication to the important part recombination took in the evolutionary process of shaping these genmes: four times more sites have been affected by recombination-driven substitutions than have been affected by mutation-driven substitutions. Estimates of population mutation rate ($\theta$), recombination rate ($\rho$), and average tract length ($\delta$) were obtained for the different alignment blocks analyzed (Figure S\ref{fig:scatter3}). Global estimates of these three population parameters were obtained by a length-weighted average (Materials and Methods). The local estimates typically seem to be moderately fluctuating around the global average, with the exception of what apears to be a recombination hotspot at a region 200 kilobase pairs from the origin of replication of SDE1.

The species tree and estimated branch lengths, together with the global estimates of the three population parameters are used in the second phase of ClonalOrigin, where recombinant trees are sampled from the posterior distriution. One way to summarize the posterior sample of recombinant trees provided by ClonalOrigin is to observe the topologies of the local genealogies implied by it. Out of the 105 possible rooted labeled topologies spanning five taxa, the dominant topology (sampled 66.7\% of the time) was that of the species tree (Materials and Methods; Table S\ref{tab:Gene-tree-topologies}). The second most frequently sampled topology (sampled 9.3\% of the time) was the one in which SPY and SDE are sister clades (with SDD as an outgroup). Other topologies were sampled with frequency less than 4\%. This gives initial indication to the extent of recombining gene transfer  between SPY and SDE.

To gain a more fine-scale view of the main types of recombination scenarios sampled by ClonalOrigin, we looked at the recombinant edges present in the sampled recombinant trees. Following the approach of \cite{Didelot2010}, we classified these edges by types defined by distinct pairs of donor and recepient branches of the species tree (Fig. [ADD FIG 1]). For each edge type, we recorded the average number of edges of that type present in the sampled recombinant tree, and computed the ratio of this observed count and the prior expected count as implied by the species tree (considering branch lengths) and the three population parameters estimated in the preliminary round of the analysis. The base 2 logarithms of these ratios are depicted in the heatmap of Figure \ref{fig:Heatmap-of-recombination} (see Tables
S\ref{tab:heatmap} and S\ref{tab:obsheatmap}).

Among the possible recombinant edge types, we distinguished between ones which preserve the species tree topology, and ones which alter it. Our main focus was on the topology-altering types, both because our main objective was to study cross-clade gene transfer, and because our simulation study indicates that the observed counts of topology-preserving recombinant edges are somewhat less reliable {\bf[IG: POINT TO RELEVANT SECTION]}. The heatmap allows us to find recombination scenarios that deviate from the expectation dictated by the probabilistic model of ClonalOrigin. {\bf [IG: CHECK THIS FOLLOWING LINE OF ARGUMENT]} We identify two main potential causes for such a deviation: (1) reduced or increased opportunities for species to interact and exchange genetic material; (2) selection acting on newly transferred genes.

We observe a reduced count of recombinant edges sampled between SDD and the three branches of the SPY clade ($\sim$25\% of the expected count). This is consistent with the fact that these species have reduced opportunities to interact, SPY being a human pathogen and SDD being strictly veterinary. On the contrary, the number of sampled recombinant edges between SDD and branches in the SDE clade seems to be consistent with the prior, with a slight excess for edges from SDE1 and SDE2 to SDD {\bf [IG: IS THIS WORTH DWELLNG ON?]}. The greatest deviation of observed counts from expected ones was observed for recombinant edges between the SPY and SDE clades. This is consistent with the fact that these two species share an ecological niche (SDE being an opportunistic human pathogen). Interestingly, there appears to be much higher enrichment (up to four times more than expected) for recombinant edges originating in the SPY clade and ending up in the SDE clade. In a follow-up simulation study we show that this deviation from the expected count is indeed highly significant, and so is the observed asymmetry in pattern of recombination {\bf[IG: POINT TO RELEVANT SECTION]}. This bias in directionality of recombination pattern between SPY and SDE is especially intriguing in light of the hypothesis that SDE acquired human pathogenicity via gene transfer from SPY.

\subsection{Parsimony-based reconciliation of gene trees with the species tree}

{\bf[IG: DID NOT TOUCH THE FIRST 3 PARAGRAPHS. REQUIRES ``CONDENSING".]}

{\bf[IG: START WITH SOME GENERAL STATISTICS - NUMBER OF GENES, ETC?]}

{\bf [IG: MODIFIED THIS PARAGRAPH]}

As in the Bayesian analysis of ClonalOrigin, we attempted to see what types of transfer events (horizontal and recombining) were most common in our inference. We classified events into types defined by donor-recepient branches in the species tree. We recorded the number of inferred events of each type in Figures \ref{fig:hgt-heatmap} \&S\ref{fig:mowgli-recomb-heatmap} {\bf [IG: CONSIDER MOVING BOTH INTO SUPPLEMENT]}. Unlike in the Bayesian analysis, we cannot normalize these counts by their expectation, since we do not have any model-based expectation for them. This restrict our ability to interprest these counts. We did notice, however, a very high number of inferred recombining gene transfer events from the ancestral SPY branch to the ancestral SDE branch, consistent with our observations from the Bayesian analysis. With horizontal gene transfer, we do not see a similar pattern, but we do observe a high number of horizontal transfers from the ancesral SDE branch to SDD. {\bf [IG: IS THERE ANYTHING ELSE WE CAN SAY ABOUT THIS? MAYBE SHOW A SCATTER PLOT OF THE TWO TYPES OF EVENTS? ]}

\subsection{Gene categories associated with recombining and horizontal gene transfer}

%\subsection{Association of virulence genes and horizontal gene transfer}

We set out to seek cateories of genes that are significantly enriched for transfer events of either type (recombining and  horizontal). For horizontal transfer, we used a score based on the number of events inferred by the parsinomy-based method, and for recombining gene transfer, we used both the parsimony-based inferred counts and a recombination intensity score based on the Bayesian analysis (Materials and Methods). First, we checked association of transfers with virulence genes {\bf [IG: A SENTENCE ON WHERE WE TAKE THESE GENES FROM]}. This set of virulence genes was found to be significantly associated with horizontal gene transfers (P-value $<1.33 \times 10^{-11}$), but not with recombining gene transfers (P-value $>0.90$ for parsimony-based score and P-value $>??$ for recombnation intensity). We also found a marginally significant association between virulence genes and gene duplications (P-value $<7.13 \times 10^{-3}$).

%\subsection{Transporters, transferases, aminoacyl tRNA synthetases, and transposases}

{\bf [IG: I THINK THE FOLLOWING 3 PARAGRAPHS SHOULD BE MOVED TO DISCUSSION]}

 Molecular evolutionary forces could act on genetic materials of organisms, and changes to the genotypes might lead to some phenotypes that could change biological functions of the species. So, the bacterial gene transfer process as an evolutionary force could be associated with some biological functions. We
attempted to find which biological functions were associated with recombining or
horizontal gene transfers.

Because bacteria as a unicellular organism should be able to respond to changes to environment, some evolutionary pressures must have been acting on genes encoding membrane-bound molecules. Recombining gene transfer process seemed to be more associated with biological functions that had been critical of sensing queues of bacterial environmental changes.  Sugar phosphotransferase classes
were ranked top in the list; phosphrylation could be important in signal
transduction to recognize environmental changes.  The transport functional
categories of carbohydrate transport and sodium ion transport also belonged to
genes encoding membrane-bound molecules.  Also highly ranked were fatty acid and
lipid biosyntheses that must be involved in producing membrane-bound molecules.
Horizontal gene transfer process would allow bacteria to acquire foreign genes
without the requirement of homologous regions in the bacteria.  Transposases
would help foreign genes be excised from a donor genome and integrated into a
recipient genome.

In addition to these biological categories, we also found several aminoacyl tRNA synthetases (GO Term translation GO:0006412) in association with recombining gene transfer.  In the parsimony-based approach the term translation was top
ranked (Table S\ref{tab:go-events-recombining}).  They had been a focus of gene
transfer in bacterial species \citep{Woese2000}.

{\bf [IG: DO WE NEED/WANT THE FOLLOWING DISCUSSION? MAYBE IN MATERIALS AND METHODS - JUSTIFYING WHY WE MEASURE INTENSITY USING TOPOLOGY-CHANGING REC EDGES.]}

 Some of the functional categories were unexpected; functional categories concerning ribosomes were found to be associated with recombining gene transfer.  Because genes related with ribosomes were considered to be housekeeping genes, gene transfer in which
would be rare, it was unexpected that ribosome related genes were found to be
associated with recombining gene transfer.  Because it was a bizarre result, we
investigated the association by considering different recombinant edge types. We
noted that two different classes of recombinant edge types: one that changes
species tree topologies, and another that does not. Because recombinant edges
that did not change species tree topology might be difficult to infer reliably,
considering those recombinant edges in computing recombination intensity might
lead to dubious results.  Ribosome related functional categories were found to
be top ranked if we took those recombinant edges that did not change species
tree topology (Tables S\ref{tab:functional-notopology} and
S\ref{tab:functional-all}).  In Table \ref{tab:functional} ribosome related
functional categories were not top ranked.  Therefore, high recombination
intensity in ribosome related genes were likely to be attributed to
topology-not-changing recombinant edges. We would be cautious of recombining
transfer of genes related with ribosome related functions.

\subsection{Comparison of Bayesian and parsimony-based inferences of recombining gene transfer}

>>>>

\section*{A list of our major and minor findings or observations}
\begin{enumerate}
\item We showed that recombining gene transfer from SPY to SDE was strong
enough.
\item A few evidences for the net directionality from SPY to SDE:
time-constrained gene transfer between SPY to SDE also showed the directionality
from SPY to SDE.
\item We showed that virulence genes were associated with horizontal gene
transfer. 
\item We found that recombining gene tranfer processes are involved in
membrane-bound protein encoding genes and aminoacyl tRNA synthetases.
\item We found that horizontal gene transfer processes are involved in
transposases.
\item We provided evidence that recombination process affected more sites along
the core genome of SPY and SDE than mutation process.
\item We found a recombination hotspot at 200 kb from the origin of replication
along the genome of SDE1.
\item We showed that species-tree-topology-changing recombinant edges were more
reliably recovered than species-tree-topology-not-changing edges.
\item We showed that recombination intensity might be useful in finding genes
under high recombining gene transfer.
\item We found that SPY and SDE are more closely related than SDD and SDE in
about 10\% of the core genome.
\item Recombination intensity and number of recombining gene transfer events
were positively correlated.
\item We found reduction of numbers of genes after divergence of SDE and SPY.
\item We confirmed the species tree that relates the five genomes.
\item We listed several virulence genes associated with recombining and
horizontal gene transfer.
\end{enumerate}

\section*{Suggestions about the manuscript}
\begin{enumerate}
\item Ilan thought that a species tree must be shown at the outset of the
section \textbf{Results}.
\item Can selections after gene transfer be tested?
\end{enumerate}

@@

Therefore, the research topic on gene exchange between SPY and SDE has been
performed in various ways. Sachse et al.\ \citet{Sachse2002} showed gene transfer from SDE to
SPY for the downstream of the genomic region of streptococcal pyogenic
exotoxins. Davies et al.\ \citet{Davies2007} investigated phage 3396 from SDE as a vehicle of
gene transfer from SDE to SPY. They \citet{Davies2009} also demonstrated that an
integrative conjugative element of SDE allowed gene transfer from SDE to other
streptococci. Ahmad et al.\ \citet{Ahmad2009} did not conclusively deduce the net
directionality of housekeeping gene transfer between SPY and SDE.  Notably,
Kalia et al.\ \citet{Kalia2001} attempted to show that the net direction of gene transfer from
SPY to SDE was more dominant than the other direction.  Because these researches
on the controversial issue, net directionality of gene transfer between SPY and
SDE, have been limited to relatively small regions in the genomes primarily on
representative loci from the Multi Locus Sequence Typing (MLST) database, it
remains to be elusive to determine the genome-wide degrees to which genomic
segments flowed between the two species, and net directions of thereof.

As sequencing costs plummet \citep{Mardis2011}, bacterial genome sequencing
produces individual genomes of high quality even within a bacterial species
\citep{Tettelin2009a}.  Studies on horizontal gene transfer in prokaryotes have
used genes in very large sets of species from a great deal of genomes available 
\citep{Koonin2001}.  Although this type of study gave us insights into deep
evolutionary histories of prokaryotes, with multiple individual genomes per
species we now can go beyond such rough broad study on gene transfer towards
fine-scaled history of gene transfer between very closely related prokaryotes.
For instance, Liu et al.\ \citet{Liu2009} studied gene transfer events in
\textit{Lactobacillus bulgaricus} and \textit{Streptococcus thermophilus} using
the compositional approach of Karlin \citet{Karlin2001}. Luo et al.\ \citet{Luo2011} investigated
gene transfer events in \textit{Eschericha Coli} using the embedded quartet
decomposition due to Zhaxybayeva et al.\ \citet{Zhaxybayeva2006}. Hamady et al.\ \citet{Hamady2006} combined
phylogenetic and compositional approaches to detecting horizontal gene transfer.
Caro-Quintero et al.\ \citet{Caro-Quintero2011} studied a number of \textit{Shewanella baltica}
genomes to identify the genetic elements that enabled the species to adapt to
redox gradients.  Although several complete genomes from SPY have been
available, Suzuki et al.\ \citet{Suzuki2011} just recently added one more complete genome of
SDE to the set of a previously unique SDE genome \citep{Shimomura2011}. These
multiple SDE genomes prompted us to study genome-wide gene flows between SPY and
SDE that afflicted human beings around the globe.

@@

Recently, a method of inferring homologous recombination with whole bacterial
genomes was developed \citep{Didelot2010}. The model of homologous recombination
allowed them to study recombining gene flows between bacteria individuals of
\textit{Bacillus cereus} species in the core genomes.   In the model, a species
tree that relates individuals genomes is augmented by attaching
\textit{recombinant edges} between species tree branches that can be used to
measure gene transfers.  Other evolutionary processes acting on bacterial
genomes include genome rearragement, translocation, inversion, duplication,
gain, loss, etc.  These processes did not readily lend themselves to
statistically more sound model-based approaches such as maximum likelihood.
Instead, a parsimony-based approach to inferring gene duplication, loss, and
transfer events was developed by Doyon et al.\ \citet{Doyon2011}.  In the
approach, gene trees are inferred for gene families that are constructed from
all of the genomes, and are reconciled with a speices tree to infer
parsimoniously events.  We noted that the first approach could be used to infer
transfer of genomic segments due to homologous recombination, and the second
could be useful in studying horizontal gene transfer.

In short, the complete genomes of the two species of SPY and SDE presented us a
unique opportunity to study gene transfer over relatively recent time scales.
We took advantage of the model-based approach for inferring homologous
recombination, and the parsimony-based one for detecting gene duplication, loss,
and horizontal transfer.  We found that virulence genes were more associated
non-recombining horizontal gene transfer than non-virulence genes.  We compared
both of the results in recombining gene transfer events to show that the
parsimony-based method inferred recombining gene transfer events in biological
functional categories similar to those inferred with the model-based approach.
This was the first attempt as far as we know of systematically analyzing SPY and
SDE in genome-wide scale to detect gene transfer events not only in the core
genome but also in the pan genome, which made the analysis complete in terms of
the amount of used genome data. Our analysis of the core genomes also showed for
the first time that the net directionality of recombining gene transfer from SPY
to SDE was stronger at the genome scale.

@@

\texttt{We show a net directionality of recombining gene transfers from SPY to SDE. 
Describe our arguments of the net directionality. 
Number of recombinant edges.
Confidence about the numbers of recombinant edges.
Not all of genes experience recombining gene transfers from SPY to SDE.
Gene transfer counts from the parsimony-based approach; the power of the
parsimony-based approach may be limited.
Is this number of recombinant edges good enough for the argument?
Reflection of the previous analyses, 
heat maps of real data analysis, 
and supporting simulations.}

\subsection*{Recombining gene transfer from SPY to SDE at genome-wide scale}
% \subsection*{Bayesian analysis of recombining gene transfer}

We followed the pipeline and methods recommended by Didelot et al.
\cite{Didelot2010} to pursue a model-based study of recombining gene transfer. A
preliminary step in the pipeline involved estimating the species
tree with branch lengths (via \texttt{ClonalFrame}), and three population
parameters of mutation rate, recombination rate, and average recombinant tract
length (via a first phase of \texttt{ClonalOrigin}).  Because 
\textit{S.\ equi} ssp.\textit{\ equi} strain 4047 (SEE)
diverged significantly from the others among the six genomes (Table S\ref{tab:genome}), we
could not obtain sufficiently long multiple sequence alignments with
it, and had to exclude it from the model-based analysis.  The estimated species tree
using the alignment of the five genomes (Figure
S\ref{fig:mauve}) had the expected topology \texttt{The figure 1 must be nicer
than one that now.} (see Figure \ref{fig:tree5}). \texttt{ClonalFrame} also
estimated
a relative strength of recombination compared to mutations ($r/m=4.66$), 
indicating potentially significant roles of recombination in the evolutionary
process of shaping these genomes; four times more sites had been affected by
recombination-driven substitutions than by mutation-driven substitutions.
($r/m$; Table S\ref{tab:clonalframe}).

Estimates of population mutation rate ($\theta$), recombination rate ($\rho$),
and average tract length ($\delta$) were obtained for the alignment blocks
analyzed (Figure S\ref{fig:scatter3}), and 
medians of each of the three parameters weighted over block lengths were
obtained as global estimates of the three population
parameters (Table S\ref{tab:three} and see \textbf{Materials and Methods}).  
The local estimates typically seemed to be
moderately fluctuating around the global average, with the exception of what
appeared to be a recombination hotspot at a region 200 kilobase pairs from the
origin of replication of SDE1.

The species tree and estimated branch lengths, together with the global
estimates of the three population parameters were used in the second phase of
ClonalOrigin, where recombinant trees were sampled from the posterior
distribution. One way to summarize the posterior sample of recombinant trees
provided by ClonalOrigin was to observe the topologies of the local genealogies
induced by it. Out of the 105 possible rooted labeled topologies spanning five
taxa, the dominant topology (sampled 66.7\% of the time) was that of the species
tree (Materials and Methods; Table S\ref{tab:Gene-tree-topologies}). The second
most frequently sampled topology (sampled 9.3\% of the time) was the one in
which SPY and SDE were sister clades (with SDD as an outgroup). Other topologies
were sampled with frequency less than 4\%. This gave initial indication to the
extent of recombining gene transfer  between SPY and SDE.

To gain a more fine-scale view of the main types of recombination scenarios
sampled by ClonalOrigin, we looked at the recombinant edges present in the
sampled recombinant trees. Following the approach of Didelot et al. \cite{Didelot2010}, we
classified these edges by types defined by distinct pairs of donor and recipient
branches of the species tree (Figure \ref{fig:tree5}). For each edge type, we recorded
the average number of edges of that type present in the sampled recombinant
tree, and computed the ratio of this observed count and the prior expected count
as implied by the species tree (considering branch lengths and their relative
positions) and the three
population parameters estimated in the preliminary round of the analysis. The
base 2 logarithms of these ratios are depicted in the heatmap of Figure
\ref{fig:Heatmap-of-recombination} (see also Tables
S\ref{tab:heatmap} and S\ref{tab:obsheatmap}).

Among the possible recombinant edge types, we distinguished between ones that
preserve the species tree topology, and ones that alter it. We focused
on the topology-altering types, both because our main objective was to study
cross-clade gene transfer, and because our simulation study indicates that the
observed counts of topology-preserving recombinant edges were 
unreliably inferred {\bf[IG: POINT TO RELEVANT SECTION]}. The heatmap allowed to find
recombination scenarios that deviated from the expectation dictated by the
probabilistic model of ClonalOrigin. {\bf [IG: CHECK THIS FOLLOWING LINE OF
ARGUMENT]} We envisioned two main potential causes for such a deviation: (1)
reduced or increased opportunities for species to interact and exchange genetic
material; (2) selection acting on newly transferred genes.

We observed a reduced count of recombinant edges sampled between SDD and the
three branches of the SPY clade ($\sim$25\% of the expected count). This is
consistent with the fact that these species have reduced opportunities to
interact with each other, SPY being a human pathogen and SDD being strictly veterinary. On the
contrary, the number of sampled recombinant edges between SDD and branches in
the SDE clade seemed to be consistent with the prior, with a slight excess for
edges from SDE1 and SDE2 to SDD {\bf [IG: IS THIS WORTH DWELLNG ON?]}. The
greatest deviation of observed counts from expected ones was observed for
recombinant edges between the SPY and SDE clades. It was concordant with the
fact that these two species share an ecological niche; SDE was an
opportunistic human pathogen. Interestingly, there appeared to be much higher
enrichment (up to four times more than expected) for recombinant edges
originating in the SPY clade and ending up in the SDE clade. In a follow-up
simulation study we show that this deviation from the expected count was indeed
highly significant, and so was the observed asymmetry in pattern of recombination
{\bf[IG: POINT TO RELEVANT SECTION]}. This bias in directionality of
recombination pattern between SPY and SDE was especially intriguing in light of
the hypothesis that SDE acquired human pathogenicity via gene transfer from SPY.

\subsection*{Parsimony-based reconciliation of gene trees with the species tree}

{\bf[IG: DID NOT TOUCH THE FIRST 3 PARAGRAPHS. REQUIRES ``CONDENSING".]}
{\bf[IG: START WITH SOME GENERAL STATISTICS - NUMBER OF GENES, ETC?]}
\texttt{SCC: Check this ``condensed'' version.}

Figure \ref{fig:Gene-duplication-loss} shows gene gain, loss, duplication, and
two different gene transfer events along the lineages of the phylogenetic tree
that relates the six genomes. SPY had less genes than the other species. All of
the genomes were used to construct 2314 gene families.  The largest family
consisted of 59 genes while nearly half of gene families (1066) had exactly one
gene per species, or six genes total (Figure S\ref{fig:famsizes}).  We estimated
a larger number of recombining gene transfer events than that of horizontal gene
transfer events; 944 and 294, respectively.  Duplications in extant lineages
were more frequent than those in the other lineages.  SDE lineages from the root
appeared to have increased in the number of genes while the common ancestor of
SPY1 and SPY2 decreased in the number of genes.  Whereas gene loss events were
either comparable to or less than gene gain ones along most of the lineages, it
was not along two internal branches of SPY and SDE.  This indicated that SPY and
SDE lineages had reduced number of genes to specialize in their specific niches
after each split from their corresponding direct ancestors.

As in the Bayesian analysis of ClonalOrigin, we attempted to see which type of
horizontal or recombining transfer events was most common in our inference.  We
classified events into types defined by donor-recipient branches in the species
tree. We recorded the number of inferred events of each type in Figures
S\ref{fig:hgt-heatmap} and S\ref{fig:mowgli-recomb-heatmap}. Unlike in the
Bayesian analysis, we could not normalize these counts by their expectation
because we did not have any model-based expectation for them. This restricted
our ability to interprest these counts. We did notice, however, a large number
of inferred recombining gene transfer events from the ancestral SPY branch to
the ancestral SDE branch, consistent with our observations from the Bayesian
analysis. With horizontal gene transfer, we did not observed a similar pattern,
but we did observe a large number of horizontal transfers from the ancesral SDE
branch to SDD.

\subsection*{Gene categories associated with recombining and horizontal gene transfer}

We set out to seek categories of genes that were significantly enriched for
recombining or horizontal transfer events. For horizontal
transfer, we used a score based on the number of events inferred by the
parsinomy-based method, and for recombining gene transfer, we used both the
parsimony-based inferred counts and a recombination intensity score based on the
Bayesian analysis (see \textbf{Materials and Methods}). First, we checked association of
transfers with virulence genes, which were provided from a previous study
of Suzuki et al.\ \citet{Suzuki2011}.
This set of virulence genes was found to be significantly associated
with horizontal gene transfers (P-value $<1.33 \times 10^{-11}$), but not with
recombining gene transfers. We also found a marginally
significant association between virulence genes and gene duplications (P-value
$<7.13 \times 10^{-3}$).

\subsection*{Comparison of Bayesian and parsimony-based inferences of recombining gene transfer}

The similarity between the two lists of gene categories found to be associated
with recombining gene transfers in the separate analyses -- Bayesian and
parsimony-based --  indirectly indicated concordance between the two independent
approaches.  In order to obtain a more direct measure of correlation, we used
the set of genes present in both analyses to compute the correlation between
recombination intensity scores and the number of recombining gene transfers
inferred by the parsimony-based approach. We observed a positive correlation of
$0.39$ (P-value $2.2\times10^{-16}$; Figure S\ref{fig:cmpcomowgli}) between the
two measures.  For comparison, the correlation between recombination intensity
and the number of horizontal gene transfers was 0.03 (P-value 0.18;
S\ref{fig:cmpcomowglihgt}).

{\bf [IG: MAYBE THERE IS A BETTER WAY TO DEMONSTRATE THE CORRELATION BETWEEN METHODS]}.

\subsection*{A closer study of virulence genes}

Often computational and statistical analyses on genomes could benefit other
researchers when those resources were provided via Genome Browser.  We wished to
provide research community especially in SPY and SDE bacteria with results from
our analysis of gene transfer.  The resource shown in Figure S\ref{fig:ucsc} for
gene transfer is available at \url{http://strep-genome.bscb.cornell.edu}.  For
each of the five individual genomes as a reference genome several annotation
tracks were displayed: alignment blocks, recombination rate per site for each
block, local gene tree topologies, recombination intensity, locations of
virulence genes, posterior probability of recombining gene transfer. The genome
track of probability of recombining gene transfer allowed us to easily browse
the bacterial genome to locate regions that we might be interested in.  Using
the genome browser tracks we located virulence genes with notable posterior
recombination probability (Table S\ref{tab:virrecomb}).  About 10\% of the genes
comprised of virulence facters, about half of which were found in the core
genome of the five genome alignment \citep{Suzuki2011}.  By sorting virulence
genes by the number of horizontal gene transfer events we listed several genes
with high ranks (Table S\ref{tab:virhgt}).
% Events counts for genes.

The process of gene transfer may or may not move whole genes from one species to
another \citep{Chan2009}. We observed that units of recombining gene
transfer could be multiple genes (Table S\ref{tab:genes-transfer}). In some
cases, parts of a gene were inferred to have been transferred, for instance,
transfer from SPY1 to SDE1 was inferred for 29\% of the aspartyl-tRNA synthetase
gene. Recombinant tracts often were shown to span two genes.  In other cases,
the inferred transfer units spaned a set of as much as eight adjacent genes:
loci from SpyM3\_1518 to SpyM3\_1525.

\section*{Discussions}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\subsection*{Distinguishing two gene transfer processes shows different functional
categories associations}

@@

% The above was moved from Results of functional category analysis.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

{\bf [IG: MOVE TO DISCUSSION:]}
A full model-based approach would be desirable in analyzing data for testing
relevant hypotheses more systematically. Model complexity might not allow easy
access to the implementation of a model. Our study involved a model-based
approach and a parsimony-based one for understanding evolution in the core
genomes of several bacteria individuals. Because of inability of the model-based
method for studying the pan genomes we had to resort to the parsimony-based
method. However, the parsimony-based method was also applied to the core
genomes. We wished to compare the two methods in their results of the core
genome analyses.  Because of two independent analyses on the same data set, this
would allow us to be more confident of the results if the two analyses performed
similarly Generally, phylogenetic methods of detecting gene transfers often take
advantage of discrepancy between a reference species tree and gene trees. The
two approaches that we employed in this study were no exception. 
Another contribution of our study to bacterial gene transfer research is to
elucidate the net directionality of gene flow from SPY to SDE in the core
genomes. Although the issue of net directionality has been tackled, our study
was the first attempt of resolving the issue in genome-wide scale. Yet, it still
remains to be seen whether this net directionality of gene flow from SPY to SDE
happened in the pan genome. A model-based method with more refined models would
be necessary for the reserach direction because the parsimony-based method we
employed simply could not show any net directionality between SPY and SDE in the
pan genome. Although we are unsure whether much larger number of genomes could
be useful in finding net directionality, we consider that more SDE genomes would
be beneficial to future studies. 
%
% The above was moved from Results of functional category analysis.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
